Reasoning About a Machine with Local Capabilities

### 2019-02-03 toplas/paper.tex
Include call in appendix and mention as part of heap based CC

\subsection{Encapsulation of Local State}
\texttt{f1} and \texttt{f2} in
\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability
machine's encapsulation of local state. They are very similar: both
store some local state, call an untrusted piece of code ($\adv$), and
then test whether the local state is unchanged. They differ in the way
they do this. Program \texttt{f1} uses our stack-based
calling convention (captured by \texttt{scall}) to call
the adversary, so it can use the available stack to store its local
state.  On the other hand, \texttt{f2} uses malloc to
allocate memory for its local state and uses a calling convention based on heap
allocated activation records (described in the technical appendix) to invoke
the adversarial code.

@@

\subsection{Macros}
Implementation of the macros used in \texttt{scall}. Implementations of the
macros not presented here can be found in the technical
appendix~\citep{technical_appendix}.\\
\label{app:macros}
\texttt{push $r$}
\begin{lstlisting}
  lea r_stk 1
  store r_stk r
\end{lstlisting}
\texttt{pop $r$}
\begin{lstlisting}
  load r r_stk
  minus r_t1 0 1
  lea r_stk r_t1
\end{lstlisting}
\texttt{rclear $r_1,\dots, r_n$}
\begin{lstlisting}
  move $r_1$ 0
  move $r_2$ 0
  $\dots$
  move $r_n$ 0
\end{lstlisting}
\texttt{mclear $r$}
\begin{lstlisting}
  move r_t $r$
  getb r_t1 r_t
  geta r_t2 r_t
  minus r_t2 r_t1 r_t2
  lea r_t r_t2
  gete r_t2
  minus r_t1 r_t2 r_t1
  plus r_t1 r_t1 1
  move r_t2 pc
  lea r_t2 $\var{ofc}_\var{end}$ 
  move r_t3 pc
  lea r_t3 $\var{ofc}_\var{iter}$
iter:
  jnz r_t2 r_t1
  store r_t 0
  lea r_t 1
  plus r_t1 r_t1 1
  jmp r_t3
end:
  move r_t 0
  move r_t1 0
  move r_t2 0
  move r_t3 0
\end{lstlisting}
Where $\var{ofc}_\var{end}$ and $\var{ofc}_\var{iter}$ are the offsets to the
label \texttt{end} % 9?
and \texttt{iter}, % 2?
respectively.

>>>>

\subsection{Encapsulation of Local State}
\texttt{f1} and \texttt{f2} in The programs in \figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's encapsulation of local state.
They are very similar: both store some local state, call an untrusted piece of code ($\adv$), and then test whether the local state is unchanged.
They differ in the way they do this.
Program \texttt{f1} uses our stack-based calling convention (captured by \texttt{scall}) to call the adversary, so it can use the available stack to store its local state.
On the other hand, \texttt{f2} uses malloc to allocate memory for its local state and uses \texttt{call} a calling convention based on heap allocated activation records (described in Appendix~\ref{app:macros}) to invoke the adversarial code.

@@

\subsection{Macros}
\label{app:macros}
Implementation of the macros used in \texttt{scall}. Implementations of the
macros not presented here can be found in the technical
appendix~\citep{technical_appendix}.\\
\texttt{push $r$}
\begin{lstlisting}[xleftmargin=0.4cm]
  lea r_stk 1
  store r_stk r
\end{lstlisting}
\texttt{pop $r$}
\begin{lstlisting}[xleftmargin=0.4cm]
  load r r_stk
  minus r_t1 0 1
  lea r_stk r_t1
\end{lstlisting}
\texttt{rclear $r_1,\dots, r_n$}
\begin{lstlisting}[xleftmargin=0.4cm]
  move $r_1$ 0
  move $r_2$ 0
  $\dots$
  move $r_n$ 0
\end{lstlisting}
\texttt{mclear $r$}
\begin{lstlisting}[xleftmargin=0.4cm]
  move r_t $r$
  getb r_t1 r_t
  geta r_t2 r_t
  minus r_t2 r_t1 r_t2
  lea r_t r_t2
  gete r_t2
  minus r_t1 r_t2 r_t1
  plus r_t1 r_t1 1
  move r_t2 pc
  lea r_t2 $\var{ofc}_\var{end}$ 
  move r_t3 pc
  lea r_t3 $\var{ofc}_\var{iter}$
iter:
  jnz r_t2 r_t1
  store r_t 0
  lea r_t 1
  plus r_t1 r_t1 1
  jmp r_t3
end:
  move r_t 0
  move r_t1 0
  move r_t2 0
  move r_t3 0
\end{lstlisting}
Where $\var{ofc}_\var{end}$ and $\var{ofc}_\var{iter}$ are the offsets to the label \texttt{end} % 9
and \texttt{iter}, % 2
respectively.
\newline\newline
\texttt{call} $r(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$
\newline
The \texttt{call} macro constitutes a calling convention based on heap allocated activation records.
This alternative to \texttt{scall} is included to illustrate that the logical relation can be used to reason about other calling conventions.
In the following, $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
\begin{itemize}
\item Set up activation record
\item Create local enter capability for activation (protected return pointer)
\item Clear unused registers
\item Jump
\item Upon return: Run activation code
  \begin{itemize}
  \item Restore private registers
  \item Jump to return capability
  \end{itemize}
\end{itemize}
\begin{lstlisting}[xleftmargin=0.4cm]
  malloc r_t $size$ 
// store private state in activation record
  store r_t r_priv,1
  lea r_t 1
  store r_t r_priv,2
  lea r_t 1
  ...
  lea r_t 1
  store r_t r_priv,n
  lea r_t 1
// store old pc
  move r_t1 pc
  lea r_t1 $\var{off}_\var{end}$ 
  store r_t r_t1
  lea r_t1 1
// store activation record
  store r_t encode(i_1)
  lea r_t1 1
  ...
  lea r_t1 1  
  store r_t encode(i_m)
  lea r_t1 $k$ 
  restrict r_t1 encodePermPair(($\local$,e))
  move r_0 r_t1
// Clear unused registers
  rclear R // R = RegisterName - {r,pc,r_0,r_args}
  jmp r
end:
\end{lstlisting}
Where $\bar{r_{\var{priv}}} = r_{\var{priv},1}, \dots, r_{\var{priv},n}$, $size$ is the size of the activation record, $\var{off}_\var{end}$ is the offset to the \texttt{end} label, and $k$ is $m-1$, i.e. the offset to the first instruction of the activation code.\newline
The activation record. The instructions correspond to $i_1,\dots,i_m$ in the above.
\begin{lstlisting}[xleftmargin=0.4cm]
  move r_t pc
  getb r_t1 r_t
  geta r_t2 r_t
  minus r_t1 r_t1 r_t2
// load private state
  lea r_t r_t1
  load r_priv,1 r_t
  lea r_t 1
  load r_priv,2 r_t
  lea r_t 1
  ...
  lea r_t 1
  load r_priv,n r_t
  lea r_t 1
// load old pc
  load pc r_t
\end{lstlisting}
